---
title: "README"
author: "Melinda K. Higgins, PhD."
date: "September 27, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# N736Fall2017_HELPdataset

N736 Fall 2017 - Details and Supporting Files and Code for HELP dataset

**Data - Basic Data - No Formatting - No Codebook**

* `help.csv` is the RAW basic file with all of the data. This file does NOT have any labels or "codes" defined; i.e. not "formatting" applied [https://github.com/melindahiggins2000/N736Fall2017_HELPdataset/blob/master/help.csv](https://github.com/melindahiggins2000/N736Fall2017_HELPdataset/blob/master/help.csv)

* `help.RData` is the basic data file in a binary format readable by `R`; you can load this into `R` by simply running `load("help.RData")`. This "loads" the data.frame `helpdata` into your Global Environment in `R`. [https://github.com/melindahiggins2000/N736Fall2017_HELPdataset/blob/master/help.Rdata](https://github.com/melindahiggins2000/N736Fall2017_HELPdataset/blob/master/help.Rdata)

* `help.sav` is the basic data file in SPSS format. However, this does NOT have any labelling - there is no codebook for this version of the data. [https://github.com/melindahiggins2000/N736Fall2017_HELPdataset/blob/master/help.sav](https://github.com/melindahiggins2000/N736Fall2017_HELPdataset/blob/master/help.sav)

**Data - With Formatting - With a Codebook**

* `helpmkh.sav` is the SPSS formatted datafile WITH the codebook applied - all of the items have labels and where applicable the responses have labels also (i.e. the "codes" and "labels"). [https://github.com/melindahiggins2000/N736Fall2017_HELPdataset/blob/master/helpmkh.sav](https://github.com/melindahiggins2000/N736Fall2017_HELPdataset/blob/master/helpmkh.sav)

* `helpmkh.RData` is the dataset WITH Labels applied in a binary format readable by `R`; you can load this into `R` by simply running `load("helpmkh.RData")`. This "loads" the data.frame `help.spss` into your Global Environment in `R`. The object `help.spss` indicates that this data was read in from the SPSS formatted file WITH the codebook applied - thus the `attributes()` for the variables in this dataset will have labels and "factors" with codes and labels applied. [https://github.com/melindahiggins2000/N736Fall2017_HELPdataset/blob/master/helpmkh.RData](https://github.com/melindahiggins2000/N736Fall2017_HELPdataset/blob/master/helpmkh.RData)

* `helpmkh.sas7bdat` along with the SAS program `helpmkh.sas` creates the `formats.sas7bcat` formats catalog. Together these help you "apply" the codebook formats to the dataset in `SAS`.
    - [https://github.com/melindahiggins2000/N736Fall2017_HELPdataset/blob/master/helpmkh.sas7bdat](https://github.com/melindahiggins2000/N736Fall2017_HELPdataset/blob/master/helpmkh.sas7bdat)
    - [https://github.com/melindahiggins2000/N736Fall2017_HELPdataset/blob/master/helpmkh.sas](https://github.com/melindahiggins2000/N736Fall2017_HELPdataset/blob/master/helpmkh.sas)
    - [https://github.com/melindahiggins2000/N736Fall2017_HELPdataset/blob/master/formats.sas7bcat](https://github.com/melindahiggins2000/N736Fall2017_HELPdataset/blob/master/formats.sas7bcat)

**`R` and `SAS` CODE to get you started**

* `loadHELPdata.R` - use this R script to load the HELP dataset with the codebook applied - this has the variable labels and coded levels [https://github.com/melindahiggins2000/N736Fall2017_HELPdataset/blob/master/loadHELPdata.R](https://github.com/melindahiggins2000/N736Fall2017_HELPdataset/blob/master/loadHELPdata.R)

* `loadHELPdata.sas` - run this SAS program AFTER running `helpmkh.sas` to get the dataset into your WORK library. [https://github.com/melindahiggins2000/N736Fall2017_HELPdataset/blob/master/loadHELPdata.sas](https://github.com/melindahiggins2000/N736Fall2017_HELPdataset/blob/master/loadHELPdata.sas)

## Variables in HELP dataset 

### Variable List and Item Labels

All variables are listed below except for `racegrp` and `substance` which are character type variables. These are listed in the next section below. For this table, all variable names, labels and minimum and maximum values are listed (excluding missing data `NA`s).

```{r, echo=FALSE, message=FALSE, warning=FALSE}
load("helpmkh.RData")
helpdata <- help.spss
library(tidyverse)

getlist <- function(x){
  data.frame(attributes(x)$label,
             min(x, na.rm=TRUE),
             max(x, na.rm=TRUE))
}

dat1 <- helpdata %>%
  select(-racegrp, -substance)

namesdat1 <- names(dat1)

library(purrr)
ldf <- purrr::map_df(dat1, getlist)
row.names(ldf) <- namesdat1

knitr::kable(ldf,
             digits=1,
             col.names = c("Item Label","Minimum","Maximum"),
             caption="HELP Variables without `racegrp` and `substance`")

```




```{r, echo=FALSE, message=FALSE, warning=FALSE}
#load("helpmkh.RData")
#helpdata <- help.spss
#library(tidyverse)

# create a function to get the label
# label output from the attributes() function
#getlabel <- function(x) attributes(x)$label
#a <- getlabel(helpdata$age)

#library(purrr)
#ldf <- purrr::map_df(helpdata, getlabel) # this is a 1x15 tibble data.frame
# t(ldf) # transpose for easier reading to a 15x1 single column list

# using knitr to get a table of these
# variable names for Rmarkdown
#library(knitr)
#knitr::kable(t(ldf),
#             col.names = c("Variable Label"),
#             caption="HELP Variables")
```

## Variables with Coded Levels and Their Labels

These next set of tables provide all the variables that have coded levels and labels (i.e. the "codebook" and values used for `PROC FORMAT` in `SAS` and treated as "factors" in `R`).

### Codes for `racegrp` and `substance`

```{r, echo=FALSE, message=FALSE, warning=FALSE}

knitr::kable(names(table(helpdata$racegrp)),
             col.names = attributes(helpdata$racegrp)$label)

knitr::kable(names(table(helpdata$substance)),
             col.names = attributes(helpdata$substance)$label)
```

### Codes for rest of the variables ...

```{r, echo=FALSE, message=FALSE, warning=FALSE}
codetable <- function(x,captitle="  "){
  knitr::kable(attributes(x)$labels,
             col.names = attributes(x)$label,
             caption=captitle)
  }

codetable(helpdata$treat,"HELP Variable `treat`")

codedvars <- helpdata %>%
  select(treat, female, homeless, g1b, 
         satreat, drinkstatus, anysubstatus,
         linkstatus, f1a, f1b, f1c, f1d, f1e,
         f1f, f1g, f1h, f1i, f1j, f1k, f1l,
         f1m, f1n, f1o, f1p, f1q, f1r, f1s, f1t)

#sapply(codedvars, codetable)

codetable(helpdata$female,"HELP Variable `female`")
codetable(helpdata$homeless,"HELP Variable `homeless`")
codetable(helpdata$g1b,"HELP Variable `g1b`")
codetable(helpdata$satreat,"HELP Variable `satreat`")
codetable(helpdata$drinkstatus,"HELP Variable `drinkstatus`")
codetable(helpdata$anysubstatus,"HELP Variable `anysubstatus`")       
codetable(helpdata$linkstatus,"HELP Variable `linkstatus`")
codetable(helpdata$f1a,"HELP Variable `f1a`")
codetable(helpdata$f1b,"HELP Variable `f1b`")
codetable(helpdata$f1c,"HELP Variable `f1c`")
codetable(helpdata$f1d,"HELP Variable `f1d`")
codetable(helpdata$f1e,"HELP Variable `f1e`")        
codetable(helpdata$f1f,"HELP Variable `f1f`")
codetable(helpdata$f1g,"HELP Variable `f1g`")
codetable(helpdata$f1h,"HELP Variable `f1h`")
codetable(helpdata$f1i,"HELP Variable `f1i`")
codetable(helpdata$flj,"HELP Variable `flj`")
codetable(helpdata$f1k,"HELP Variable `f1k`")
codetable(helpdata$f1l,"HELP Variable `f1l`")         
codetable(helpdata$f1m,"HELP Variable `f1m`")
codetable(helpdata$f1n,"HELP Variable `f1n`")
codetable(helpdata$f1o,"HELP Variable `f1o`")
codetable(helpdata$f1p,"HELP Variable `f1p`")
codetable(helpdata$f1q,"HELP Variable `f1q`")
codetable(helpdata$f1r,"HELP Variable `f1r`")
codetable(helpdata$f1s,"HELP Variable `f1s`")
codetable(helpdata$f1t,"HELP Variable `f1t`")
```

